### 2013-06-28 Publications/2013-TMI/03-introduction.tex
Revised intro
Close #81. I also add the pdf containing @meribach comments on the
intro with the fixes.

\section{Introduction}
\label{sec:introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%
\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which recently 
has accounted for a growing interest in its application to structural 
connectivity analyses of the brain at the macro-scale
\citep{craddock_imaging_2013}. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
for the clinical application \citep{griffa_structural_2013} as well.

Early \gls{dwi} research mainly focused on the improvements of imaging 
methodologies by a better understanding of the diffusion effect, and improving
image reconstruction methodologies. Currently, the connectome extraction 
and analysis relies on a large chain of sophisticated computational techniques
including acquisition, reconstruction, modeling and model fitting, image processing, 
fibre tracking, connectivity mapping, visualization, group studies, and inference
\citep{daducci_connectome_2012}. This growing complexity has given rise to 
challenging issues towards reliable structural information about the neuronal tracts 
presented in, and also difficulties on the statistical analysis 
\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
regarding the impact of the different sources of inaccuracy and variability along
the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
cheng_characteristics_2012,jones_white_2013} are scarce.
Here,we shall address three tasks included within  the image processing stage in 
a unified approach: brain tissue segmentation in diffusion space 
(\autoref{sec:dwi_segmentation}), 
correction of geometrical distortions (\autoref{sec:distortion}), and structural 
image registration to diffusion coordinate space (\autoref{sec:registration}). 
These tasks are generally
solved independently, or combined in pairs. However, there exist fundamental 
coupling relationships that can be exploited to obtain a simultaneous solution to 
the three problems. This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

@@

\Gls{epi} are known to suffer from geometrical distortions and artifacts
due to, mainly, three sources: the subject motion in between acquiring 
different sampling directions, the induced \emph{Eddy currents} on the scanner 
coils and finally distortions caused by the magnetic susceptibility inhomogeneity
present at tissue interface. \emph{Susceptibility distortions} happen along the 
phase-encoding direction, and are most appreciable in the frontal part of 
the brain due to the air/tissue interface located around the sinuses.
Two implications are associated to this artifact: the signal loss caused on
highly distorted regions, and a significant added difficulty in 
registering with structural images (e.g. T1-weighted).

@@

None of the presented methods have claimed for definite results, mainly due to the lack of 
a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
regions, or do not provide sub-pixel resolution results
\citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
(\autoref{sec:registration}) are more compelling, minimizing the activity on this line which is 
currently being considered to be included in reconstruction algorithms 
\citep{kumazawa_improvement_2013}. \emph{Golden}-standard evaluation frameworks have been 
proposed for the segmentation validation \citep{jha_task-based_2012} on the task of lesion 
detection in visceral organs.  \todo[inline]{reference to a summary table}

@@

The last addressed task in the image processing stage of the studied 
connectivity analysis pipelines is the structural information co-registration.
The need for this registration step is mainly raised when defining the 
nodes of the connectivity matrix. This definition step is named
\textit{cortical parcellation} as it imposes the regions that will be
considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
surface. The common procedure to accomplish this clustering is using
a predefined parcellation in a high-resolution structural T1-weighted 
atlas (e.g. \citep{greve_accurate_2009}). The second goal that has 
justified the applicability of structural
images co-registration is susceptibility distortion correction
as discussed in \autoref{sec:distortion}. In any case, susceptibility 
distortion hinders a rigid-registration solution to this problem.
Additionally, when applying non-linear intensity-based
registration algorithms other difficulties have to be addressed as
the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the
inherent inadequacy of the B0 contrast (the only \textit{directionally
independent} existing in the raw \gls{dwi} data) due to the almost null
difference in intensity between \gls{wm} and \gls{gm} pixels. Early 
registration methods appeared targeting the distortion correction.
They quickly standardized the choice of T2 as anatomically correct source
to be registered against the B0 in \gls{dwi}. B0 images have a very similar
contrast to T2 weighted due to the parallelism of acquisition sequences.
However, parcellation is defined in T1 weighted space, adding an additional 
registration step is generally overlooked by existing literature. Even though this
processing stage has a very low complexity compared to the remaining tasks of 
the pipeline process, it is one additional source of inconsistency
and unreliability.

>>>>

\section{Introduction}
\label{sec:introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%
\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which 
accounts for a growing interest in its application to structural 
connectivity analyses of the brain at the macro-scale
\citep{craddock_imaging_2013}. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
for the clinical application \citep{griffa_structural_2013} as well.

Early \gls{dwi} research mainly focused on the improvements of imaging 
methodologies by a better understanding of the diffusion effect, and improving
image reconstruction methodologies. Currently, the connectome extraction 
and analysis relies on a large chain of sophisticated computational techniques
including acquisition, reconstruction, modeling and model fitting, image segmentation
and registration, fibre tracking, connectivity mapping, visualization, statistical 
group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
has given rise to challenging issues towards reliable structural information about 
the neuronal tracts presented in, and also difficulties on the statistical analysis 
\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
regarding the impact of the different sources of inaccuracy and variability along
the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
cheng_characteristics_2012,jones_white_2013} are scarce.
Here, we shall address three tasks included within  the image processing stage in 
a unified approach: brain tissue segmentation in diffusion space 
(\autoref{sec:dwi_segmentation}), 
correction of geometrical distortions (\autoref{sec:distortion}), and structural 
image registration to diffusion coordinate space (\autoref{sec:registration}). 
These tasks are generally
solved independently, or combined in pairs. However, there exist fundamental 
coupling relationships that can be exploited to obtain a simultaneous solution to 
the three problems. This joint approach satisfactory impacts the downstream
outcomes of the whole pipeline with the increase of the internal consistency.

@@

\Gls{epi} are known to suffer from geometrical distortions and artifacts
due to, mainly, three sources: the subject motion in between acquiring 
different sampling directions, the induced \emph{Eddy currents} on the scanner 
coils, and, finally, distortions caused by the magnetic susceptibility inhomogeneity
present at tissue interface. \emph{Susceptibility distortions} happen along the 
phase-encoding direction, and are most appreciable in the frontal part of 
the brain due to the air/tissue interface located around the sinuses.
Two implications are associated to this artifact: the signal loss caused on
highly distorted regions, and a significant added difficulty in 
registering with structural images (e.g. T1-weighted).

@@

None of the presented methods have claimed for definite results, mainly due to the lack of 
a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
regions, or do not provide sub-pixel resolution results
\citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
(\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi} 
segmentation. However, recently \citep{kumazawa_improvement_2013} proposed to include this
task within reconstruction algorithms. Finally, some \emph{golden}-standard evaluation
frameworks for task-based validation have been proposed for the segmentation results
(e.g. lesion detection in visceral organs \citep{jha_task-based_2012}).
\todo[inline]{reference to a summary table}

@@

The last addressed task in the image processing stage of the studied 
connectivity analysis pipelines is the structural information co-registration.
The need for this registration step is mainly raised when defining the 
nodes of the connectivity matrix. This definition step is named
\textit{cortical parcellation} as it imposes the regions that will be
considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
surface. The common procedure to accomplish this clustering is using
a predefined parcellation in a high-resolution structural T1-weighted 
atlas (e.g. \citep{greve_accurate_2009}). The second goal that has 
justified the applicability of structural
images co-registration is susceptibility distortion correction
as presented in \autoref{sec:distortion}. In any case, susceptibility 
distortion hinders a rigid-registration solution to this problem.
Additionally, when applying non-linear intensity-based
registration algorithms, other difficulties have to be addressed as
the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the
inherent inadequacy of the B0 contrast (the only \textit{directionally
independent} existing in the raw \gls{dwi} data) due to the almost null
difference in intensity between \gls{wm} and \gls{gm} pixels. Early 
registration methods appeared targeting the distortion correction.
They quickly standardized the choice of T2 as anatomically correct source
to be registered against the B0 in \gls{dwi}. B0 images have a very similar
contrast to T2 weighted due to the parallelism of acquisition sequences.
However, parcellation is defined in T1 weighted space, adding an additional 
registration step is generally overlooked by existing literature. Even though this
processing stage has a very low complexity compared to the remaining tasks of 
the pipeline process, it is one additional source of inconsistency
and unreliability.
